Bad forecast, good decision: what makes a model relevant for determining policy

### 2020-10-26 manuscript/appendix.Rmd
:pen:

## Decision Context

and the manager seeks to maximize the sum of the utility derived from such a harvest and such a state, $U(X_t,H_t)$, over all time, subject to discount rate $\delta$:

@@

```{r prob_scores, message = FALSE}
## Probability scores instead of proper scores
scores <- sims %>%
  mutate(next_state = dplyr::lead(state_index), model = as.integer(model)) %>%
  filter(model < 3) %>%
  rowwise() %>%
  mutate(prob =  transitions[[model]][state_index, next_state, action_index] ) %>% 
  group_by(reps, model) %>%
  summarise(score = sum(log(prob)))

score_table <- scores %>% group_by(model) %>% summarise(mean_score = mean(score))

scores %>% mutate(model = as.factor(model)) %>% 
  ggplot(aes(x = score, group = model, fill = model)) +
  geom_histogram(binwidth = 70)

@@

>>>>

---
title: 'Appendix for: "Bad Forecast, Good Decision..."'
date: "`r Sys.Date()`"
author: Carl Boettiger
journal: Appendix
layout: 3p
bibliography: refs.bib
#output:
#  tufte::tufte_handout:
#    dev: cairo_pdf
#    latex_engine: xelatex
output: 
  hrbrthemes::ipsum_pdf:
    dev: cairo_pdf
    latex_engine: xelatex

---

```{r graphics_setup, message=FALSE, warning=FALSE, include = FALSE}
## Plotting themes, colors, and fonts 
## aesthetics choices only, all can be omitted without consequence
library(ggthemes)
library(hrbrthemes)
library(ggplot2)
library(Cairo)
library(extrafont)
library(patchwork)
library(styler)
extrafont::loadfonts(quiet = TRUE)
ggplot2::theme_set(hrbrthemes::theme_ipsum_rc())

scale_colour_discrete <- function(...) ggthemes::scale_colour_solarized()
scale_fill_discrete <- function(...) ggthemes::scale_fill_solarized()
pal <- ggthemes::solarized_pal()(6)
txtcolor <- "#586e75"

knitr::opts_chunk$set(cache=FALSE, tidy = "styler", message = FALSE, warning = FALSE, echo = TRUE)

Here we provide annotated code necessary to completely reproduce all of the analysis
presented in the main paper.
This analysis is run in R [@R] uses `MDPtoolbox` [@MDPtoolbox] for solving 
Markov Decision Processes (MDP) using stochastic dynamic programming functionality,
`expm` for matrix exponentials [@expm], and a few custom MDP functions provided
by our package, `mdplearning` [@mdplearning].  We will also use `tidyverse` 
packages for basic manipulation and plotting [@tidyverse]. This file is also
available as an RMarkdown document [@rmarkdown] at 
<https://github.com/cboettig/bad-forecast-good-decision>

```{r setup, message=FALSE}
library(tidyverse)
library(MDPtoolbox)
library(expm)
# remotes::install_github("boettiger-lab/mdplearning")
library(mdplearning)
```

Our decision problem is defined as follows: 
A manager seeks to maximize the sum of the utility derived from such a harvest
and such a state, $U(X_t,H_t)$, over all time, subject to discount rate $\delta$:

@@

To solve the decision problem using SDP, we define
the state space and the action space on a discrete grid of 240 points.  The maximum
state is set well above the largest carrying capacity used in the models,
which limits the influence of boundary effects introduced by the transformation to
a discrete, finite grid.  Available harvest actions match the state space, effectively 
allowing any harvest level to be possible.

```{r model_definitions}
states <- seq(0,24, length.out = 240)
actions <- states
obs <- states
```

The utility (reward) of an action is set to the (realized) harvest (e.g.
a fixed price per unit fish, with no cost applied to harvest effort).  Future
utility is discounted by fixed factor of $0.99$. Classic maximum sustainable 
yield models [@Schaefer1954] ignore discounting, while modern economic optimization
models insist on it, so a small discount conforms to the latter while reasonably
approximating the former.  The qualitative conclusion is not sensitive to the 
discounting rate.

```{r}
reward_fn <- function(x,h) pmin(x,h)
discount <- 0.99
```

Alternate reward functions with varying price and cost structures are possible but do not qualitatively
impact the conclusions.  This reward function is a limiting case of any more 
complex reward, and corresponds both to classic work that does not model utility
explicitly [e.g. MSY theory, @Schaefer1954], as well as explicit assumptions
typically made in more recent models [@Reed1979]. Moreover, using a simple reward function
makes it clear that model that performs best is not doing so merely because of particular
features baked into the a carefully chosen reward rule.

# Ecological Models

The manager chooses between two different logistic growth models, each of which
can be thought of as an approximation to the underlying "true" population model:

\begin{equation}
f_i(Y) = Y + Y r_i \left( 1 - \frac{Y}{K_i} \right)
\end{equation}

Model 1 has $r_1 = 2$, $K = 16$, and $\sigma_1 = 0.05$.
Model 2 has $r_2 = 0.5, K_2 = 10$, and $\sigma_2 = 0.075$.

Meanwhile, the true population growth rate is simulated using a function with
non-linear per-capita growth:

\begin{equation}
f_i(Y) = Y + Y^4 r_i \left( 1 - \frac{Y}{K_i} \right)
\end{equation}

with $r_3 = 0.002$, $K_3 = 10$ and $\sigma_3 = 0.05$:

```{r}
# K is at twice max of f3; 8 * K_3 / 5
f1 <- function(x, h = 0, r = 2, K = 10 * 8 / 5){
  s <- pmax(x - h, 0)
  s + s * (r * (1 - s / K) )
}
f2 <- function(x, h = 0, r = 0.5, K = 10){
  s <- pmax(x - h, 0)
  s + s * (r * (1 - s / K) )
}

# max is at 4 * K / 5 
f3  <- function(x, h = 0, r = .002, K = 10){
  s <- pmax(x - h, 0)
  s + s ^ 4 * r * (1 - s / K)
}

## gather models together, indicate true model
sigma_g <- 0.05
models <- list(f1 = f1, f2 = f2, f3 = f3)
model_sigmas <- c(sigma_g, 1.5 * sigma_g, sigma_g)
true_model <- "f3"
```

On a discrete grid of possible states and actions, we can define the growth rate of a given state $X_t$ subject to harvest $H_t$,
$f(X_t,H_t)$ as set of matrices.  Each matrix $i$ gives the transition probabilities for any current state to any future state, 
given that action $i$ is taken.

```{r transition_matrices}
transition_matrices <- function(f, states, actions, sigma_g){
  n_s <- length(states)
  n_a <- length(actions)
  transition <- array(0, dim = c(n_s, n_s, n_a))
  for (k in 1:n_s) {
    for (i in 1:n_a) {
      nextpop <- f(states[k], actions[i])
      if(nextpop <= 0){
        transition[k, , i] <- c(1, rep(0, n_s - 1))
      } else if(sigma_g > 0){
        x <- dlnorm(states, log(nextpop), sdlog = sigma_g)
        if(sum(x) == 0){ ## nextpop is computationally zero
          transition[k, , i] <- c(1, rep(0, n_s - 1))
        } else {
          x <- x / sum(x) # normalize evenly
          transition[k, , i] <- x
        }
      }
    }
  }
  transition
}
```

This follows the standard setup for standard stochastic dynamic programming, see @Marescot2013.
Having defined a function to compute the transition matrix, we can use it to create matrices corresponding to each of the three models:

```{r}
transitions <- lapply(seq_along(models), 
                      function(i) transition_matrices(models[[i]], 
                                                      states, 
                                                      actions, 
                                                      model_sigmas[[i]]))
names(transitions) <- c("f1", "f2", "f3")

@@

Likewise, a corresponding matrix defining the rewards associated with each state $X$ and each harvest action $H$ can also be defined.

````{r}
## Compute reward matrix (shared across all models)
n_s <- length(states)
n_a <- length(actions)
reward <- array(0, dim = c(n_s, n_a))
for (k in 1:n_s) {
  for (i in 1:n_a) {
    reward[k, i] <- reward_fn(states[k], actions[i])
  }
}
```

# Optimal control solutions

We use value iteration to solve the stochastic dynamic program [@Marescot2013; @MDPtoolbox] for each model.  
This determines the optimal harvest policy for each possible state, given each model. Because this step is the
most computationally intensive routine, we cache the results using memosization conditioned on the transition
matrices [@memoise].  Running this code with alternate transition matrices automatically invalidates that
cache, reducing the risk of loading spurious results.

```{r sdp, results="hide"}
mdp <- memoise::memoise(mdp_value_iteration,
                        cache = memoise::cache_filesystem("cache/"))

policies <- 
  map_dfr(transitions, 
          function(P){
    soln <- mdp(P, reward, discount = discount, 
                epsilon = 0.01, max_iter = 1000, V0 = rep(0,dim(P)[[1]]))
    escapement <- states - actions[soln$policy]
    tibble(states, policy = soln$policy, escapement)
    }, 
  .id = "model")
```

# Simulations and step-ahead forecasts

We simulate fishing dynamics under the optimal policy for each model, using a simple
helper function from the `mdplearning` package. Because growth dynamics are stochastic,
we perform 100 simulations of each model from identical starting condition to
ensure results are not the result of chance alone.

```{r simulations}
library(mdplearning)
Tmax <- 100
x0 <- which.min(abs(states - 6))
reps <- 100
set.seed(12345)

## Simulate each policy reps times, with `f3` as the true model:
simulate_policy <- function(i, policy){
  mdp_planning(transitions[[true_model]], reward, discount,
           policy = policy, x0 = x0, Tmax = Tmax) %>%
    select(value, state_index = state, time, action_index = action)  %>% 
    mutate(state = states[state_index])
}

sims <- 
  map_dfr(names(transitions), 
          function(m){
            policy <- policies %>% filter(model == m) %>% pull(policy)
            map_dfr(1:reps, simulate_policy, policy = policy, .id = "reps")
          },
          .id = "model"
         )
```

Using the transition matrices directly, we can examine what each model would 
have forecast the future stock size to be in the following year when no fishing
occurs (note that for each model, we use the transition matrix that corresponds
to 'no fishing', `model[[state_index, ,1]]`) (Fig 1a, main text).

The transition matrices give the full (discretized) probability distribution,
from which we can easily calculate both the expected value and the 95% 
confidence interval.

```{r stepahead_unfished}
stepahead_unfished <- sims
stepahead_unfished$state_index <- rep(sims$state_index[sims$model == "1"],3)

stepahead_unfished <- stepahead_unfished  %>% 
  filter(model != "3") %>% 
  mutate(next_state = dplyr::lead(state_index), model = as.integer(model)) %>%
  rowwise() %>%
  mutate(expected = transitions[[model]][state_index, , 1]  %*% states,
         var = transitions[[model]][state_index, , 1]  %*% states ^ 2 - expected ^ 2,
         low = states[max(which(cumsum(transitions[[model]][state_index,,1]) < 0.025)) ],
         high = states[min(which(cumsum(transitions[[model]][state_index,,1]) > 0.975)) ],
         true = states[next_state ],
         model = paste0("f", model)) 
```

We also look at the forecast each model makes when implementing the corresponding
optimal harvest:

```{r stepahead_fished}
stepahead_fished <- sims %>% 
  filter(model != "3") %>%
  mutate(next_state = dplyr::lead(state_index), model = as.integer(model)) %>%
  rowwise() %>%
  mutate(prob =  transitions[[model]][state_index, next_state, action_index],
         expected = transitions[[model]][state_index, , action_index]  %*% states,
         var = transitions[[model]][state_index, , action_index]  %*% states ^ 2 - expected ^ 2,
         low = states[max(which(cumsum(transitions[[model]][state_index,,action_index]) < 0.025)) ],
         high = states[min(which(cumsum(transitions[[model]][state_index,,action_index]) > 0.975)) ],
         true = states[next_state],
         model = paste0("f", model)) %>%
 select(time, model, true, expected, low, high, var, prob, reps)
```

# Proper scores

It is straight forward to apply the proper scoring formula of @Gneiting2007 
based on the first two moments of the distribution to score the respective
forecasts under both the unfished and actively managed scenario sfor each model:

```{r proper_scores, message = FALSE}
# Gneiting & Raferty (2007), eq24
scoring_fn <- function(x, mu, sigma){ -(mu - x )^2 / sigma^2  - log(sigma)}

stepahead_unfished <- stepahead_unfished %>%
  mutate(sd = sqrt(var),
         score = scoring_fn(expected, true, sd))

stepahead_fished <- stepahead_fished %>%
  mutate(sd = sqrt(var),
         score = scoring_fn(expected, true, sd))
```

```{r}
predictions <- 
  stepahead_unfished %>% 
  select(time, model, reps, expected, low, high, true, score) %>% 
  mutate(scenario = "A_unfished") %>% 
  bind_rows(stepahead_fished  %>% 
          select(time, model, reps, expected, low, high, true, score) %>% 
          mutate(scenario = "B_fished")
  ) 
```

# Plotting

Having computed a single data frame of simulation data under both optimally fished
(according to each model) and unfished scenarios for each model, along with
the proper scores for the step-ahead forecasts a la @Gneiting2007, we have all 
the results necessary to generate the figures presented in the main paper.

```{r fig1a}
fig1ab <- predictions %>%
  filter(reps == "2", time < 10) %>%
  ggplot(aes(time, col = model, fill = model)) + 
  geom_point(aes(y = expected)) + 
  geom_errorbar(aes(ymin = low, ymax = high)) +
  geom_point(aes(y = true), pch = "*", size = 12, 
             alpha = 0.6) + 
  facet_wrap(~scenario, ncol = 1, labeller =as_labeller(c(
    A_unfished = "A. Without fishing",
    B_fished = "B. Managed harvest"))) +
  ylab("stock size")

Likewise we can plot the data on proper scores associated with
each prediction of each model:

```{r}
fig1cd <- predictions %>%
  ggplot(aes(x = score, group = model, fill = model)) +
  geom_histogram(binwidth = 2, show.legend = FALSE) +
  coord_cartesian(xlim = c(-100, 1), ylim = c(0,4000)) +
  xlab("Proper score") +
  facet_wrap(~scenario, ncol = 1, labeller =as_labeller(c(
    A_unfished = "C",
    B_fished = "D"))) + 
  scale_x_continuous(breaks = c(-100, 0)) +
  theme(axis.text.y = element_blank(),
        plot.margin = margin(0, 0, 0, 0, "cm"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()
        )

## Forecast ecological and economic performance

The ecological performance of the model can easily be visualized by plotting
the results of each simulation.

```{r plot_sims}
fig2a <- 
  sims %>%
  filter(time < 25, reps < 5, model != 3) %>%
  ggplot(aes(time, state, col=model, group = interaction(model,reps))) + 
  geom_line(alpha=0.3) 
```

To plot the economic value over time, we must sum up the discounted values
at each time step, and then average over replicate simulations of each model:

```{r plot_npv}
##  Net Present Value accumulates over time
npv_df <- sims %>% 
  group_by(model, reps) %>%
  mutate(npv = cumsum(value * discount ^ time)) %>%
  group_by(time, model)  %>% 
  summarise(mean_npv = mean(npv), .groups="drop") %>% 
  arrange(model, time)

optimal <- select(filter(npv_df, model == "3"), time, mean_npv)

fig2b <- 
npv_df %>%
  filter(model != "3", time %in% seq(1,100, by = 5)) %>%
  ggplot(aes(time, mean_npv)) +
  geom_line(data = optimal, lwd = 1.5, col = "grey20") +
  geom_point(aes(col=model), size = 4, alpha = 0.8) + 
  ylab("Net present value") + xlab("time")

Plotting the model functions themselves are shown in Fig 3a:

```{r plot_models}
d <- 
  map_dfc(models, function(f) f(states) - states) %>%
  mutate(state = states)

fig3a <- 
  d %>% pivot_longer(names(models), "model") %>%
  ggplot(aes(state, value, col=model)) +
  geom_point() + 
  geom_hline(aes(yintercept = 0)) + 
  coord_cartesian(ylim = c(-5, 8), xlim = c(0,16))
```

We can easily plot the optimal policy derived from each model, as shown
in Fig3b.

```{r plot_policies}
fig3b <- policies %>%
  ggplot(aes(states, escapement, col=model, lty=model)) + 
  geom_line(lwd=2)
```

# References

### 2020-10-26 manuscript/manuscript.Rmd
:pen:

---
title: "Bad forecast, good decision: what makes a model relevant for determining policy"
date: "`r Sys.Date()`"
author:
  - name: Carl Boettiger
    email: cboettig@berkeley.edu
    affiliation: ucb
#    footnote: Corresponding Author
address:
  - code: ucb
    address: "ESPM Department, University of California, 130 Mulford Hall Berkeley, CA 94720-3114, USA"
abstract: |
    With the rapid expansion of available data, it is frequently suggested that _forecasting performance_ is the ultimate test of a scientific model [e.g. @Dietz2018].  Yet the model that makes the most accurate forecasts is not always the model that yields the best decisions.  This is no accident of statistics, even formal criteria such as strictly proper scoring rules [@Gneiting2007] cannot determine the best model for decision-making because they do not consider the context of the decision process itself.  I illustrate this paradox using a classic example of ecological decision making in fisheries management.  This example also reveals how an understanding of the decision problem can both explain the poor management outcomes of model which produces the best forecast, and also see how to avoid making such a mistake.  As forecasting becomes an increasing attractive and viable possibility for assessing ecological models and applying them to real-world decisions, I hope this illustration serves as a reminder that modeling can not be done in a vacuum, but must reflect the context in which the model will be applied.
journal: Theoretical ecology
layout: 3p
bibliography: refs.bib
output: rticles::elsevier_article
keywords: 
  - transients, 
  - optimal control, 
  - adaptive management, 
  - stochasticity, 
  - uncertainty, 
  - ecological management
preamble: |
  \journal{preprint}
  ## EditorialManager adds its own linenumbers to pdf already
  \usepackage{lineno}
  \linenumbers

@@

```{r graphics_setup, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(cache=TRUE)

## Plotting theme settings -- aesthetics only, all can be omitted without consequence
library(ggthemes)
library(hrbrthemes)
library(ggplot2)
library(Cairo)
library(extrafont)
library(patchwork)
extrafont::loadfonts()
ggplot2::theme_set(hrbrthemes::theme_ipsum_rc())
#theme_set(theme_solarized(base_size=16))

@@

```{r setup, message=FALSE}
## Required computational libraries
library(tidyverse)
library(MDPtoolbox)
library(expm)
library(mdplearning) # remotes::install_github("boettiger-lab/mdplearning")
```

```{r model_definitions}
states <- seq(0,24, length.out = 240)
actions <- states
obs <- states
sigma_g <- 0.05
reward_fn <- function(x,h) pmin(x,h)
discount <- 0.99

# K is at twice max of f3; 8 * K_3 / 5
f1 <- function(x, h = 0, r = 2, K = 10 * 8 / 5){
  s <- pmax(x - h, 0)
  s + s * (r * (1 - s / K) )
}
f2 <- function(x, h = 0, r = 0.5, K = 10){
  s <- pmax(x - h, 0)
  s + s * (r * (1 - s / K) )
}

# max is at 4 * K / 5 
f3  <- function(x, h = 0, r = .002, K = 10){
  s <- pmax(x - h, 0)
  s + s ^ 4 * r * (1 - s / K)
}

# looks even closer to f2; not used
#f4  <- function(x, h = 0, r = 4e-7, K = 10){
#  s <- pmax(x - h, 0)
#  s + s ^ 8 * r * (1 - s / K) + s * (1 - s / K)
#}

models <- list(f1 = f1, f2 = f2, f3 = f3)
model_sigmas <- c(sigma_g, 1.5 * sigma_g, sigma_g)

true_model <- "f3"
```

@@

```{r transition_matrices}
# A function to compute the transition matrices for each model:
transition_matrices <- function(f,
                      states,
                      actions,
                      sigma_g){

  n_s <- length(states)
  n_a <- length(actions)

  transition <- array(0, dim = c(n_s, n_s, n_a))
  for (k in 1:n_s) {
    for (i in 1:n_a) {
      nextpop <- f(states[k], actions[i])
      if(nextpop <= 0){
        transition[k, , i] <- c(1, rep(0, n_s - 1))
      } else if(sigma_g > 0){
        x <- dlnorm(states, log(nextpop), sdlog = sigma_g)
        if(sum(x) == 0){ ## nextpop is computationally zero
          transition[k, , i] <- c(1, rep(0, n_s - 1))
        } else {
          x <- x / sum(x) # normalize evenly
          transition[k, , i] <- x
        }
      } else {
        stop("sigma_g not > 0")
      }
      reward[k, i] <- reward_fn(states[k], actions[i])
    }
  }
  transition
}

## Compute reward matrix (shared across all models)
n_s <- length(states)
n_a <- length(actions)
reward <- array(0, dim = c(n_s, n_a))
for (k in 1:n_s) {
  for (i in 1:n_a) {
    reward[k, i] <- reward_fn(states[k], actions[i])
  }
}

## Now compute the transition matrices for each model
transitions <- lapply(seq_along(models), 
                      function(i) transition_matrices(models[[i]], 
                                                      states, 
                                                      actions, 
                                                      model_sigmas[[i]]))
names(transitions) <- c("f1", "f2", "f3")

```

```{r sdp, results="hide"}
## This is the most (only) computationally expensive code chunk. use memoization to cache
mdp <- memoise::memoise(mdp_value_iteration, cache = memoise::cache_filesystem("cache/"))

policies <- 
  map_dfr(transitions, 
          function(P){
    soln <- mdp(P, reward, discount = discount, 
                epsilon = 0.01, max_iter = 1000, V0 = rep(0,dim(P)[[1]]))
    escapement <- states - actions[soln$policy]
    tibble(states, policy = soln$policy, escapement)
    }, 
  .id = "model")
```

```{r simulations}
library(mdplearning)
Tmax <- 100
x0 <- which.min(abs(states - 6))
reps <- 100
set.seed(12345)

## Simulate each policy reps times, with `f3` as the true model:

sims <- map_dfr(names(transitions), 
                function(m){
                  policy <- policies %>% filter(model == m) %>% pull(policy)
                  map_dfr(1:reps, 
                          function(i){
                            mdp_planning(transitions[[true_model]], reward, discount,
                                     policy = policy, x0 = x0, Tmax = Tmax) %>%
                              select(value, state_index = state, time, action_index = action)  %>% 
                              mutate(state = states[state_index]) # index->value
                            },
                          .id = "reps")
                },
                .id = "model")

```

```{r stepahead_unfished}
stepahead_unfished <- sims
stepahead_unfished$state_index <- rep(sims$state_index[sims$model == "1"],3)
stepahead_unfished <- stepahead_unfished  %>% 
  filter(model != "3") %>% 
  mutate(next_state = dplyr::lead(state_index), model = as.integer(model)) %>%
  rowwise() %>%
  mutate(expected = transitions[[model]][state_index, , 1]  %*% states,
         var = transitions[[model]][state_index, , 1]  %*% states ^ 2 - expected ^ 2,
         low = states[ max(which(cumsum(transitions[[model]][state_index,,1]) < 0.025)) ],
         high = states[ min(which(cumsum(transitions[[model]][state_index,,1]) > 0.975)) ],
         true = states[ next_state ],
         model = as.character(model)) 
```

```{r fig1a}
fig1a <- stepahead_unfished %>% 
  filter(reps == 2, time < 10) %>%
  ggplot(aes(time, col = model, fill = model)) + 
  geom_point(aes(y = expected), show.legend = FALSE) + 
  geom_errorbar(aes(ymin = low, ymax = high), show.legend = FALSE) +
  geom_point(aes(y = true), pch = "*", size = 12, col = "grey40", show.legend = FALSE)

```

```{r stepahead_fished}
stepahead_fished <- sims %>% 
  filter(model != "3") %>%
  mutate(next_state = dplyr::lead(state_index), model = as.integer(model)) %>%
  rowwise() %>%
  mutate(prob =  transitions[[model]][state_index, next_state, action_index],
         expected = transitions[[model]][state_index, , action_index]  %*% states,
         var = transitions[[model]][state_index, , action_index]  %*% states ^ 2 - expected ^ 2,
         low = states[ max(which(cumsum(transitions[[model]][state_index,,action_index]) < 0.025)) ],
         high = states[ min(which(cumsum(transitions[[model]][state_index,,action_index]) > 0.975)) ],
         true = states[ next_state],
         model = as.character(model)) %>%
 select(time, model, true, expected, low, high, var, prob, reps)
```

```{r fig1b}
fig1b <- stepahead_fished  %>%  
  filter(reps == 3, time < 10) %>%
  ggplot(aes(time, col = model, fill = model)) + 
  geom_point(aes(y = true), pch = "*", size = 12) +
  geom_point(aes(y = expected)) + 
  geom_errorbar(aes(ymin = low, ymax = high))
 # geom_errorbar(aes(ymin = expected - 2 * sqrt(var), ymax=expected + 2 * sqrt(var)))

```

```{r proper_scores, message = FALSE}

scoring_fn <- function(x, mu, sigma){ -(mu - x )^2 / sigma^2  - log(sigma)}

proper_scores_unfished <- stepahead_unfished %>%
  mutate(sd = sqrt(var),
         score = scoring_fn(expected, true, sd)) %>%
  select(model, reps, time, score)

fig1c <- proper_scores_unfished %>%
  ggplot(aes(x = score, group = model, fill = model)) +
  geom_histogram(binwidth = 2, show.legend = FALSE) +
  coord_cartesian(xlim = c(-100, 1))

proper_scores_fished <- stepahead_fished %>%
  mutate(sd = sqrt(var),
         score = scoring_fn(expected, true, sd)) %>%
  select(model, reps, time, score)

fig1d <- proper_scores_fished %>%
  ggplot(aes(x = score, group = model, fill = model)) +
  geom_histogram(binwidth = 2) +
  coord_cartesian(xlim = c(-100, 1))

```

@@

```{r figure1, fig.width=7, fig.height=8}
(fig1a / fig1b) /
(fig1c + fig1d)

```

@@

```{r plot_sims}
fig2a <- 
  sims %>%
  filter(time < 25, reps < 5) %>%
  ggplot(aes(time, state, col=model, group = interaction(model,reps))) + 
  geom_line(alpha=0.3) 
```

```{r plot_npv}
##  Net Present Value accumulates over time, equivalent for models with near-identical management stategy
npv_df <- sims %>% 
  group_by(model, reps) %>%
  mutate(npv = cumsum(value * discount ^ time)) %>%
  group_by(time, model)  %>% 
  summarise(mean_npv = mean(npv)) %>% 
  arrange(model, time) %>% ungroup()

optimal <- select(filter(npv_df, model == "3"), time, mean_npv)

fig2b <- 
npv_df %>%
  filter(model != "3", time %in% seq(1,100, by = 5)) %>%
  ggplot(aes(time, mean_npv)) +
  geom_line(data = optimal, lwd = 1.5, col = "grey20") +
  geom_point(aes(col=model), size = 4, alpha = 0.8) + 
  ylab("Net present value") + xlab("time")

```
```{r figure2, message = FALSE, fig.cap = "Corresponding utility (measured as mean net present value, that is: cumulative value, discounting future values by $\\delta^t$, averaged across replicates). Note that the exepcted utility under model 1, which has the worst forecast, is nearly identical to the optimal utility achieved by managing under the correct model, 3.  The utility derived from model 2 is far smaller, despite it's overall better performance in long term forecasts."}
fig2a + fig2b
```

@@

```{r plot_models}
d <- map_dfc(models, function(f) f(states) - states) %>% mutate(state = states)

fig3a <- 
  d %>% pivot_longer(names(models), "model") %>%
  ggplot(aes(state, value, col=model)) +
  geom_point() + 
  geom_hline(aes(yintercept = 0)) + 
  coord_cartesian(ylim = c(-5, 8), xlim = c(0,16))
```

```{r plot_policies}
fig3b <- policies %>%
  ggplot(aes(states,escapement, col=model, lty=model)) + geom_line(lwd=2)
```

>>>>

---
title: "Bad forecast, good decision: what makes a model relevant for determining policy"
date: "`r Sys.Date()`"
author:
  - name: Carl Boettiger
    email: cboettig@berkeley.edu
    affiliation: ucb
#    footnote: Corresponding Author
address:
  - code: ucb
    address: "ESPM Department, University of California, 130 Mulford Hall Berkeley, CA 94720-3114, USA"
abstract: |
    With the rapid expansion of available data, it is frequently suggested that _forecasting performance_ is the ultimate test of a scientific model [e.g. @Dietz2018].  Yet the model that makes the most accurate forecasts is not always the model that yields the best decisions.  This is no accident of statistics, even formal criteria such as strictly proper scoring rules [@Gneiting2007] cannot determine the best model for decision-making because they do not consider the context of the decision process itself.  I illustrate this paradox using a classic example of ecological decision making in fisheries management.  This example also reveals how an understanding of the decision problem can both explain the poor management outcomes of model which produces the best forecast, and also see how to avoid making such a mistake.  As forecasting becomes an increasing attractive and viable possibility for assessing ecological models and applying them to real-world decisions, I hope this illustration serves as a reminder that modeling can not be done in a vacuum, but must reflect the context in which the model will be applied.
journal: TBD
layout: 3p
bibliography: refs.bib
output: 
  rticles::elsevier_article:
    dev: cairo_pdf
    keep_tex: true
keywords: 
  - transients, 
  - optimal control, 
  - adaptive management, 
  - stochasticity, 
  - uncertainty, 
  - ecological management
preamble: |
  \journal{preprint}
  \usepackage{lineno}
  \linenumbers

@@

```{r, message=FALSE, warning=FALSE, include = FALSE}
## Plotting themes, colors, and fonts 
## aesthetics choices only, all can be omitted without consequence
library(ggthemes)
library(hrbrthemes)
library(ggplot2)
library(Cairo)
library(extrafont)
library(patchwork)
library(styler)
extrafont::loadfonts(quiet = TRUE)
ggplot2::theme_set(hrbrthemes::theme_ipsum_rc())

@@

```{r, include=FALSE}
rmarkdown::render("appendix.Rmd")
knitr::opts_chunk$set(cache = TRUE, echo = FALSE, message = FALSE, warning = FALSE, dev="cairo_pdf")
```

@@

@@

```{r figure1, fig.width=7, fig.height=5, include=TRUE, message=TRUE, warning=TRUE}
fig1ab + fig1cd + 
  plot_layout(widths =  c(3,1), guides = 'collect') + 
  theme(plot.margin = margin(0, 0, .1, .1, "cm"))
```

@@

```{r figure2, message = FALSE, fig.cap = "Corresponding utility (measured as mean net present value, that is: cumulative value, discounting future values by $\\delta^t$, averaged across replicates). Note that the exepcted utility under model 1, which has the worst forecast, is nearly identical to the optimal utility achieved by managing under the correct model, 3.  The utility derived from model 2 is far smaller, despite it's overall better performance in long term forecasts."}
fig2a + fig2b
```

@@
